---
title: "hw3"
output: github_document
date: "2025-10-12"
---

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Import dataset
```{r}
data("instacart")
```

There are `r nrow(instacart)` rows and `r ncol(instacart)` in the `instacart` dataset. 
The dataset include these variables: `r colnames(instacart)`.

Here are some of the key important variables with illustrative observations showing the order of each product, product information, whether it's reordered and the days since the prior order:

```{r}
set.seed(1234)
instacart |> 
  select(order_id, reordered, days_since_prior_order, product_name, aisle, department) |> 
  relocate(order_id, product_name) |> 
  slice_sample(n = 10)
```

Each row in the dataset represents a single product (`product_id`) purchase history from a user (`user_id`). The same `order_id` means that these products are purchased together as a single order. Other key information include where the product is from: `aisle_id`, `aisle`, `department` and `department_id`, and when the order happened: `order_hour_of_day`, `order_dow` (day of the week). 

Number of aisles + which have most items ordered:

```{r}
num_aisle = 
  instacart |> 
  distinct(aisle_id) |> # find distinct aisle id to get how many aisles in the dataset
  nrow()

most_items_aisle = 
  instacart |> 
  count(aisle, sort = TRUE) |> # The most items ordered aisle means that it occurs the most
  slice(1:5) |> 
  pull(aisle)
```

There are `r num_aisle` aisles, and `r most_items_aisle` have the most items ordered.

Creating the aisle plot:

```{r}
aisle_plot = 
  instacart |> 
  group_by(aisle) |> 
  summarize(
    n_items = n()
  ) |> 
  filter(n_items > 10000) |> # only select aisles with more than 10000 items
  mutate(
    aisle = fct_reorder(aisle, n_items, .desc = TRUE)
  ) |> 
  ggplot(aes(x = aisle, y = n_items)) + 
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "aisle",
    y = "number of items sold"
  ) 

aisle_plot
```

Make a table showing popular items in these three aisles: "baking ingredients", "dog food care", "packaged vegetables fruits":

```{r}
top3_table = 
  instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle, product_name) |> 
  summarize(
    times_ordered  = n(),
    .groups = "drop_last" # keep grouping as aisle but not product_name
  ) |> 
  slice_max(times_ordered, n = 3) |> 
  knitr::kable()

top3_table
```

Make a mean hour table:

```{r, message = FALSE}
mean_hour_table = 
  instacart |> 
  filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream")) |> 
  mutate(dow = factor(
    order_dow, levels = 0:6,
    labels = c("Sun","Mon","Tue","Wed","Thu","Fri","Sat") # rename the dow
  )) |>
  group_by(product_name, dow) |> 
  summarize(
    mean_hour = round(mean(order_hour_of_day, na.rm = TRUE), 2)
  ) |>
  pivot_wider(
    names_from = dow, 
    values_from = mean_hour) |>
 knitr::kable()

mean_hour_table
```


## Problem 2

```{r}
nyc_zillow_df = 
  read_csv("data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv") |> 
  pivot_longer(
    -(RegionID:CountyName),
    names_to = "month",
    values_to = "price") |> 
  janitor::clean_names() |> 
  rename(zip_code = region_name) |> 
  mutate(
    month = as_date(month),
    year = floor_date(month, unit = "year"),
    zip_code = as.numeric(zip_code)) |> 
  select(-county_name)

zip_codes_df = 
  read_csv("data/Zip Codes.csv") |> 
  janitor::clean_names() |> 
  filter(
    !(zip_code == 10463 & county == "New York"),
    !(zip_code == 11201 & county == "New York"))  |> 
  mutate(
    borough = case_match(
      county,
      "Bronx" ~ "Bronx",
      "Kings" ~ "Brooklyn",
      "New York" ~ "Manhattan",
      "Queens" ~ "Queens",
      "Richmond" ~ "Staten Island"
    )
  ) |> 
  select(zip_code, borough, neighborhood)

nyc_price_df = 
  left_join(nyc_zillow_df, zip_codes_df, by = "zip_code")
```

ZIP codes observed 116 times and fewer than 10 times:

```{r}
zipcode_obs =
  nyc_price_df |>
  filter(
    month >= ymd("2015-01-01"),
    month <= ymd("2024-08-31"),
    !is.na(price)
  ) |>
  group_by(zip_code) |>
  summarize(
    n_months = n()
  )
 
n_116_times = 
  zipcode_obs |> 
  filter(n_months == 116) |> 
  nrow()

n_fewer10 = 
  zipcode_obs |> 
  filter(n_months < 10) |> 
  nrow()

```

There are `r n_116_times` zipcodes observed 116 times and `r n_fewer10` zipcodes observed fewer than 10 times. This is because some of the zip code areas are standard residential areas but some are not.

Create a table showing the average rental price in each borough and year:

```{r}
rental_price_avg = 
  nyc_price_df |> 
  group_by(borough, year) |> 
  summarize(
    avg_price_year = round(mean(price, na.rm = TRUE), 2)
  ) |> 
  pivot_wider(
    names_from = year,
    values_from = avg_price_year
  ) |> 
  knitr::kable()

rental_price_avg
```




Make a plot showing NYC Rental Prices within ZIP codes for all available years:

```{r}
```


