---
title: "hw3"
output: github_document
date: "2025-10-12"
---

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Import dataset
```{r}
data("instacart")
```

There are `r nrow(instacart)` rows and `r ncol(instacart)` in the `instacart` dataset. 
The dataset include these variables: `r colnames(instacart)`.

Here are some of the key important variables with illustrative observations showing the order of each product, product information, whether it's reordered and the days since the prior order:

```{r}
set.seed(1234)
instacart |> 
  select(order_id, reordered, days_since_prior_order, product_name, aisle, department) |> 
  relocate(order_id, product_name) |> 
  slice_sample(n = 10)
```

Each row in the dataset represents a single product (`product_id`) purchase history from a user (`user_id`). The same `order_id` means that these products are purchased together as a single order. Other key information include where the product is from: `aisle_id`, `aisle`, `department` and `department_id`, and when the order happened: `order_hour_of_day`, `order_dow` (day of the week). 

Number of aisles + which have most items ordered:

```{r}
num_aisle = 
  instacart |> 
  distinct(aisle_id) |> # find distinct aisle id to get how many aisles in the dataset
  nrow()

most_items_aisle = 
  instacart |> 
  count(aisle, sort = TRUE) |> # The most items ordered aisle means that it occurs the most
  slice(1:5) |> 
  pull(aisle)
```

There are `r num_aisle` aisles, and `r most_items_aisle` have the most items ordered.

Creating the aisle plot:

```{r}
aisle_plot = 
  instacart |> 
  group_by(aisle) |> 
  summarize(
    n_items = n()
  ) |> 
  filter(n_items > 10000) |> # only select aisles with more than 10000 items
  mutate(
    aisle = fct_reorder(aisle, n_items, .desc = TRUE)
  ) |> 
  ggplot(aes(x = aisle, y = n_items)) + 
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "aisle",
    y = "number of items sold"
  ) 

aisle_plot
```


