---
title: "hw3"
output: github_document
date: "2025-10-12"
---

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Import dataset
```{r}
data("instacart")
```

There are `r nrow(instacart)` rows and `r ncol(instacart)` in the `instacart` dataset. 
The dataset include these variables: `r colnames(instacart)`.

Here are some of the key important variables with illustrative observations showing the order of each product, product information, whether it's reordered and the days since the prior order:

```{r}
set.seed(1234)
instacart |> 
  select(order_id, reordered, days_since_prior_order, product_name, aisle, department) |> 
  relocate(order_id, product_name) |> 
  slice_sample(n = 10) |> 
  knitr::kable()
```

Each row in the dataset represents a single product (`product_id`) purchase history from a user (`user_id`). The same `order_id` means that these products are purchased together as a single order. Other key information include where the product is from: `aisle_id`, `aisle`, `department` and `department_id`, and when the order happened: `order_hour_of_day`, `order_dow` (day of the week). 

Number of aisles + which have most items ordered:

```{r}
num_aisle = 
  instacart |> 
  distinct(aisle_id) |> # find distinct aisle id to get how many aisles in the dataset
  nrow()

most_items_aisle = 
  instacart |> 
  count(aisle, sort = TRUE) |> # The most items ordered aisle means that it occurs the most
  slice(1:5) |> 
  pull(aisle)
```

There are `r num_aisle` aisles, and `r most_items_aisle` have the most items ordered.

Creating the aisle plot:

```{r}
  instacart |> 
  group_by(aisle) |> 
  summarize(
    n_items = n()
  ) |> 
  filter(n_items > 10000) |> # only select aisles with more than 10000 items
  mutate(
    aisle = fct_reorder(aisle, n_items, .desc = TRUE)
  ) |> 
  ggplot(aes(x = aisle, y = n_items)) + 
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "aisle",
    y = "number of items sold"
  ) 

```


Make a table showing popular items in these three aisles: "baking ingredients", "dog food care", "packaged vegetables fruits":

```{r}
  instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle, product_name) |> 
  summarize(
    times_ordered  = n(),
    .groups = "drop_last" # keep grouping as aisle but not product_name
  ) |> 
  slice_max(times_ordered, n = 3) |> 
  knitr::kable()
```

Make a mean hour table:

```{r, message = FALSE}
  instacart |> 
  filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream")) |> 
  mutate(dow = factor(
    order_dow, levels = 0:6,
    labels = c("Sun","Mon","Tue","Wed","Thu","Fri","Sat") # rename the dow
  )) |>
  group_by(product_name, dow) |> 
  summarize(
    mean_hour = round(mean(order_hour_of_day, na.rm = TRUE), 2)
  ) |>
  pivot_wider(
    names_from = dow, 
    values_from = mean_hour) |>
 knitr::kable()
```

## Problem 2

Import the zillow and zipcode datset and merge:

```{r, message = FALSE}
nyc_zillow_df = 
  read_csv("data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv") |> 
  pivot_longer(
    -(RegionID:CountyName),
    names_to = "month",
    values_to = "price") |> 
  janitor::clean_names() |> 
  rename(zip_code = region_name) |> 
  mutate(
    month = as_date(month),
    year = floor_date(month, unit = "year"),
    zip_code = as.numeric(zip_code)) |> 
  select(-county_name)

zip_codes_df = 
  read_csv("data/Zip Codes.csv") |> 
  janitor::clean_names() |> 
  filter(
    !(zip_code == 10463 & county == "New York"),
    !(zip_code == 11201 & county == "New York"))  |> 
  mutate(
    borough = case_match(
      county,
      "Bronx" ~ "Bronx",
      "Kings" ~ "Brooklyn",
      "New York" ~ "Manhattan",
      "Queens" ~ "Queens",
      "Richmond" ~ "Staten Island"
    )
  ) |> 
  select(zip_code, borough, neighborhood)

nyc_price_df = 
  left_join(nyc_zillow_df, zip_codes_df, by = "zip_code")
```

ZIP codes observed 116 times and fewer than 10 times:

```{r}
zipcode_obs =
  nyc_price_df |>
  filter(
    month >= ymd("2015-01-01"),
    month <= ymd("2024-08-31"),
    !is.na(price)
  ) |>
  group_by(zip_code) |>
  summarize(
    n_months = n()
  )
 
n_116_times = 
  zipcode_obs |> 
  filter(n_months == 116) |> 
  nrow()

n_fewer10 = 
  zipcode_obs |> 
  filter(n_months < 10) |> 
  nrow()

```

There are `r n_116_times` zipcodes observed 116 times and `r n_fewer10` zipcodes observed fewer than 10 times. This is because some of the zip code areas are standard residential areas but some are not.

Create a table showing the average rental price in each borough and year:

```{r, message = FALSE}
  nyc_price_df |> 
  mutate(year = format(year, "%Y")) |>
  group_by(borough, year) |> 
  summarize(
    avg_price_year = round(mean(price, na.rm = TRUE), 2)
  ) |> 
  pivot_wider(
    names_from = year,
    values_from = avg_price_year
  ) |> 
  knitr::kable()
```

Based on the table, average rental prices increase slowly each year (excluding the impact of COVID-19), but after the pandemic, average rental prices increase much higher than before in each year.

Make a plot showing NYC Rental Prices within ZIP codes for all available years:

```{r, message = FALSE}
rental_price_plot = 
  nyc_price_df |> 
  ggplot(aes(x = month, y = price, group = zip_code, color = borough)) +
  geom_line(alpha = 0.5, linewidth = 0.4) +
  facet_wrap(~borough) +
  labs(
    x = "year",
    y = "price",
    title = "NYC Rental Prices by Borough"
  ) +
  theme_minimal() + 
  theme(
    legend.position = "none"
  )

rental_price_plot
```

Based on the plot, it is clear that rental prices in all the five boroughs increase over time. From the plot, it is also clear that there is a drop in price during the pandemic, and in part of 2024, which is more significant in Queens, Brooklyn, and Manhattan.

Make a plot showing  ZIP-code-level rental prices across borough in 2023:

```{r, message = FALSE}
price_2023_avg_plot = 
  nyc_price_df |> 
  filter(year(month) == 2023, !is.na(price)) |>
  group_by(borough, zip_code) |>
  summarize(
    avg_price = mean(price, na.rm = TRUE)) |> 
  ggplot(aes(x = borough, y = avg_price, color = borough)) +
  geom_boxplot() +
  labs(
    title = "2023 Average Rental Prices by Borough"
  ) + 
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank()
  )

price_2023_avg_plot
```

Based on the plot, Manhattan has the highest overall rental prices in 2023 and Staten Island has the lowest. The price range in Staten Island is also the smallest.

Export the two plots:
```{r, message = FALSE}
library(patchwork)
result_plots = rental_price_plot + price_2023_avg_plot
ggsave("results/combined_plot.png", result_plots, width = 12, height = 10, bg = "white")
```

## Problem 3

Import the two datasets, tidy and merge them

```{r, message = FALSE}
accel_df = 
  read_csv("data/nhanes_accel.csv") |> 
  drop_na() |> 
  janitor::clean_names()

demo_df = 
  read_csv("data/nhanes_covar.csv", skip = 4) |> 
  drop_na() |> 
  filter(age >= 21) |> 
  janitor::clean_names() |> 
  mutate(
    sex = factor(
      sex, 
      levels = c(1,2), 
      labels = c("male", "female")),
    education = factor(
      education,
      levels = c(1, 2, 3),
      labels = c("Less than high school", "High school equivalent", "More than high school")
    )
  )

accel_demo_df = 
  left_join(demo_df,accel_df, by = "seqn") 
```

Create table for education:

```{r, message = FALSE}
demo_df |> 
  group_by(education, sex) |> 
  summarize(n = n()) |> 
  pivot_wider(
    names_from = education,
    values_from = n
  ) |> 
  knitr::kable()

demo_df |> 
  ggplot(aes(x = age, fill = sex)) +
  geom_histogram() + 
  facet_wrap(~education, ncol = 1) + 
  labs(
    title = "Age Distribution by Education Level and Sex"
  )
```

Based on the education table by sex and the age distribution by education level and sex, there is not much difference on the education level between the two sexes. More than high school education level is the most common education level among the participants.


Aggregate and get the plot:

```{r}
accel_demo_df |> 
  pivot_longer(
    cols = starts_with("min"),
    names_to = "minute_times",
    values_to = "mins"
  ) |> 
  group_by(seqn, age, sex, education) |> 
  summarize(
    total_act = sum(mins, na.rm = TRUE)
  ) |> 
  ggplot(aes(x = age, y = total_act, color = sex)) + 
  geom_point() + 
  geom_smooth(se = FALSE) + 
  facet_wrap(~education, nrow = 1) + 
  labs(
    title = "Total daily accelerometer activity vs age by education"
  ) + 
  theme_minimal()
```

Based on the plot, we can see that as people age, the total daily activity decreases. From the smooth line, participants whose educational level are less than high school do not show differences in sex. For participants who have high school or more educational level, females tend to have more total daily activity.


```{r}
accel_demo_df |> 
    pivot_longer(
    cols = starts_with("min"),
    names_to = "minute_times",
    values_to = "mins"
  ) |> 
  mutate(
    minute_num = as.numeric(str_remove(minute_times, "min")),
    hour = floor((minute_num - 1) / 60) # keep hour starts from 0 to 23 ...
  ) |> 
  group_by(sex, education, hour) |> 
  summarize(
    mean_act_hour = mean(mins, na.rm = TRUE)
  ) |> 
  ggplot(aes(x = hour, y = mean_act_hour, color = sex)) + 
  geom_point() + 
  geom_smooth(se = FALSE) + 
  facet_grid(~education) + 
  labs(
    title = "24 Hour Activity Patterns by Education Level and Sex", 
    x = "hour of day",
    y = "average activity"
  ) + 
  theme_minimal()
  
```

Based on this graph, we can conclude that the peak activity happens at noon (peak at 12, and high from 10 to 15). Looking at the raw data, we also see that activity starts to decline at around 18:00 to 19:00 and starts decrease from 5:00. In addition, the mean activity per hour does not show significant differences between the three educational levels, but women who have high school or more educational level tends to have higher activity (which is also shown in the graph above). 

